Forecasting the 2024 US Presidential Election

### 2024-11-04 paper/paper.qmd
Minor fix in introduction section
Updated the README and made consistency changes in the paper

---
title: "Forecasting the 2024 US Presidential Election"
subtitle: "Analyzing Demographic Patterns and Predicting Swing States"
author: 
  - Tina Kim
  - David Flores
  - Kevin Shao
thanks: "Code and data are available at: https://github.com/DavidFJ207/USPresidentialForecast"
date: today
date-format: long
abstract: This paper analyzes recent swing state polling data to forecast outcomes for the 2024 U.S. Presidential Election. Our findings indicate a close race, with Republicans leading in key states such as Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats hold a slight edge in Arizona. The national polling average also leans towards Republican, suggesting a Republican win in the 2024 Election. These insights give us a clearer picture of voter preferences across states and highlight the importance of targeted campaign strategies. 
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper addresses a gap in election forecasting by using recent polling data from key swing states to provide a prediction of the 2024 election outcome. We aim to determine which candidate—Kamala Harris or Donald Trump—is more likely to win this election. Our response variable is in turn the probability of one candidate winning the election based on our chosen Emerson poll data. When one candidate’s support hovers just below 50%, it becomes challenging to predict their election as the 47th President of the United States, especially considering the potential influence of non-participating voters. To increase the accuracy of our predictions, we will develop two models: one estimating the probability of Harris, representing the Democratic Party, winning the election, and the other estimating the probability of Trump, representing the Republican Party, winning. By comparing these predictions, we aim to arrive at a clearer conclusion about who is more likely to be elected as the next President of the United States.

In our models, we identified eight pivotal swing states—Georgia, Michigan, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin—crucial for predicting the 2024 election outcome. We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, and current leanings toward Trump or Harris. Our election forecast resulted in 6 out of 8 pivotal states, as well as the national polling average, leaning towards the Republican party.

@@

From this analysis, we identified eight critical swing states: Georgia, Michigan, National, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin. These states will be central to our election prediction. Although Minnesota emerged as a potential swing state, we excluded it due to the wide range of polling data, as explained in @tbl-final-summary. Additionally, we will consider the national average in our overall prediction.

>>>>

---
title: "Forecasting the 2024 US Presidential Election"
subtitle: "Analyzing Demographic Patterns and Predicting Swing States"
author: 
  - Tina Kim
  - David Flores
  - Kevin Shao
thanks: "Code and data are available at: https://github.com/DavidFJ207/USPresidentialForecast"
date: today
date-format: long
abstract: This paper analyzes recent swing state polling data to forecast outcomes for the 2024 U.S. Presidential Election. Our findings indicate a close race, with Republicans leading in key states such as Georgia, Nevada, North Carolina, New Hampshire, Wisconsin, and Pennsylvania, while Democrats hold a slight edge in Michigan and Arizona. The national polling average also leans towards Republican, suggesting a Republican win in the 2024 Election. These insights give us a clearer picture of voter preferences across states and highlight the importance of targeted campaign strategies. 
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper uses recent polling data from key swing states to provide a prediction of the 2024 election outcome. We aim to determine which candidate—Kamala Harris or Donald Trump—is more likely to win this election. Our response variable is in turn the probability of one candidate winning the election based on our chosen Emerson poll data. When one candidate’s support hovers just below 50%, it becomes challenging to predict their election as the 47th President of the United States, especially considering the potential influence of non-participating voters. To increase the accuracy of our predictions, we will develop two models: one estimating the probability of Harris, representing the Democratic Party, winning the election, and the other estimating the probability of Trump, representing the Republican Party, winning. By comparing these predictions, we aim to arrive at a clearer conclusion about who is more likely to be elected as the next President of the United States.

In our models, we identified eight pivotal swing states—Michigan, Georgia, Nevada, New Hampshire, North Carolina, Pennsylvania, Wisconsin, and Arizona—crucial for predicting the 2024 election outcome. We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, and current leanings toward Trump or Harris. Our election forecast resulted in 6 out of 8 pivotal states, as well as the national polling average, leaning towards the Republican party.

@@

From this analysis, we identified eight critical swing states: Michigan, Georgia, Nevada, New Hampshire, North Carolina, Pennsylvania, Wisconsin, and Arizona. These states will be central to our election prediction. Although Minnesota emerged as a potential swing state, we excluded it due to the wide range of polling data, as explained in @tbl-final-summary. Additionally, we will consider the national average in our overall prediction.
